---
title: "Introduction to reporttool"
author: "Kristian D. Olsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This package is meant for a particular use, which is to work with data from
surveys to generate batch reports. To do so, we need to keep track of the labels
(questions) for each variable, in addition to latent associations for use in structural
modelling (PLS-PM).

To achieve this, `reporttool` provides a `Survey` class which is built using R6,
and stuffs some extra information into private (hidden) `fields` inside it. For users,
a `Survey` should look and behave pretty much like a normal `data.frame`, `data.table`
or `tbl` (dplyr). Internally `Survey`'s helps us keep track of the hidden fields, toward
the end goal of automating tables, plots and reports.

## Creating a Survey

We can create a new survey with the generic function `survey()`, which creates a
`Survey` based on the class of the supplied data. All `Survey` objects, also have
an additional subclass which reflects the underlying data. If we call `survey()` on 
a `data.table`, the subclass is `Survey_dt`. To coerce a `data.frame` to a 
`Survey_dt`, we can call `survey_dt()` directly.

Note: Because `Survey` is an R6 class, I have opted to use a capital **S** for
the class name, and use a lowercase letters for the constructor functions.


Read in sample data:
```{r}
sav <- haven::read_sav(system.file("extdata", "sample.sav", package = "reporttoolDT"))
```

To create a `Survey`, we can call `survey()` on the data:
```{r}
require(reporttoolDT)
tbl <- survey(sav)
class(tbl)
```

As we can see, simply calling `survey()` on the data will create a `Survey_tbl`,
this is because `haven` returns a `tbl_df` by default. If we instead wanted a 
`Survey_dt` (i.e., use `data.table`), we have two options:

```{r}
dtm <- survey(data.table::as.data.table(sav)) # Option 1: Manual.
dt <- survey_dt(sav) # Option 2: Use the Survey_dt constructor.
"Survey_dt" %in% class(dt) && "Survey_dt" %in% class(dtm)
```

That's it. We have 3 subclasses for `Survey`, these are `Survey_df`, `Survey_dt`
and `Survey_tbl`. We can either call the generic constructor `survey()`,or a specific 
one which coerces the input data, these are: `survey_df()`, `survey_dt()` and
`survey_tbl()` (requires dplyr).

## Working with surveys

When working with surveys, we can use `[` and `[[` just like on a regular `data.frame`:
```{r}
df <- survey_df(sav)
df[["test"]] <- "test"
df[1L, "test"]
```

Or `data.table`: 
```{r}
dt <- survey_dt(sav)
head(dt[, test := "test"][1L, test])
```

When subsetting a `Survey`, the new object will also be a `Survey`:
```{r}
dt_slice <- dt[1:5L, .(test)]
class(dt_slice)
```

But only if the operation returns an object that inherits from `data.frame`:
```{r}
class(dt_slice[, test])
```

Overall, a `Survey` should behave like the underlying data. The only exception to
this, is that `$` is reserved for accessing public fields in R6 objects:
```{r}
identical(df$test, df[["test"]]) # FALSE
```

Because the `Survey` class is implemented in R6, they are mutable and allows the 
`Survey_dt` subclass to have hidden fields behave similar to the `data.table` itself,
which will be shown in the next section.

