---
title: "Other functions (reporttoolDT)"
author: "Kristian D. Olsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Other functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Foreword

`reporttoolDT` contains several "convenience"-functions that are made to make ones life easier when working with Survey data. This vignette attempts to provide a list with these functions and a small example to showcase their usage - we'll use the same dataset as in previous vignettes:

```{r, warning = FALSE, message = FALSE}
require(dplyr)
require(reporttoolDT)

sav <- read_data(system.file("extdata", "raw_data.sav", package = "reporttoolDT"))
srv <- survey_tbl(sav) %>%
  set_association(common = TRUE) %>%
  set_config(name = "Example", segment = "B2C", cutoff = .3) %>%
  set_marketshare(CompanyA = .3, CompanyB = .5, CompanyC = .2) %>%
  set_translation(language = "english") %>%
  latents_mean() 
```

## Adding variables

`add_latent_spread()` adds all spread variables to the data: 

```{r}
out <- srv %>% add_latent_spread()
out %>% select(ends_with("_spread")) %>% head()
```

`add_weight()` creates a weight ("w") based on marketshares for each entity:

```{r}
out <- srv %>% add_weight()
out %>% select(w) %>% head()
```

## Cleaning

`clean_scale()` helps you convert likert-scales stored as character vectors to numeric by extracting numbers from text:

```{r}
clean_scale(c("1 Not happy", 2:9, "10 Very happy"))
```

`rescale_100()` converts a 10-point scale to a 100-point scale:

```{r}
rescale_100(1:10)
```

`rescale_10()` does the reverse of `rescale_100`:

```{r}
rescale_10(rescale_100(1:10))
```

`str_to_numeric()` lets you convert strings in general to numerics:

```{r}
str_to_numeric("string 1 with many values 9", FUN = mean)
# Output is 5 because there are two numbers: 1+9/2
```

## Recoding

`recode()` let's you exchange one or more values for another in a vector:

```{r}
recode(c("A", "B", "C"), GroupA = c("A", "C"), GroupB = "B", factor = TRUE)
```

`spread_10()` helps you spread a 10-point likert as follows:

```{r}
spread_10(1:10)
```

`spread_100()` does the same for 100-point scales:

```{r}
spread_100(rescale_100(1:10))
```

## Tables

`latent_table()` creates a table with latent scores:

```{r}
srv %>% group_by(q1) %>% latent_table()
```

`manifest_table()` does the same for manifest variables:

```{r}
srv %>% group_by(q1) %>% manifest_table()
```

There is also `impact_table()` for outer weights (sorted by weight), when you have read a survey with argument `outer_weight = TRUE`.

## Charts

`latent_plot()` creates a plot with latent scores:

```{r, fig.height = 5, fig.width = 7}
srv %>% group_by(q1) %>% latent_plot()
```

`manifest_plot()` does the same for manifest scores:

```{r, fig.height = 5, fig.width = 7}
srv %>% group_by(q1) %>% manifest_plot()
```

You can also use `flow_chart()` after reading outer weights by specifying `inner_weight = TRUE` in `read_survey()`.

## More information

[Introduction](https://github.com/itsdalmo/reporttoolDT/blob/master/vignettes/introduction.Rmd): 
```{r, eval = FALSE}
vignette("introduction", package = "reporttoolDT")
```

[Survey-class](https://github.com/itsdalmo/reporttoolDT/blob/master/vignettes/survey.Rmd): 
```{r, eval = FALSE}
vignette("survey", package = "reporttoolDT")
```

[Preparing data](https://github.com/itsdalmo/reporttoolDT/blob/master/vignettes/prepare.Rmd):
```{r, eval = FALSE}
vignette("prepare", package = "reporttoolDT")
```
