---
title: "Survey class (reporttoolDT)"
author: "Kristian D. Olsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Survey class}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## What is a Survey?

As we saw in the introduction, the `Survey` is essentially a `data.frame` which stores some extra information about the variables and the survey as a whole. To do this, the `Survey` class is built on top of [R6](https://github.com/wch/R6) and [R6Frame](https://github.com/itsdalmo/R6Frame).

Note: Because `Survey` is an R6 class, I have opted to use a capital **S** for the class name, and use a lowercase letters for the constructor functions.

Let's start by reading in the sample data, creating a `Survey` again and looking at it's class:

```{r}
require(reporttoolDT)
sav <- read_data(system.file("extdata", "sample.sav", package = "reporttoolDT"))
srv <- survey(sav)
class(srv)
```

As we can see from the class of our `srv` object, it inherits from "R6" in addition to "Survey". When I say the `Survey` is essentially a `data.frame`, this is because the data rests inside the survey itself, and can be accessed using `$`:

```{r}
# Show data/class
head(srv$data[, 1:5L], n = 3L)
class(srv$data)
```

If we look at the class again, we see that `srv$data` is just the same data that was read in and stored in the `sav` object. The difference is that `survey()` converts from labelled variables to factors, and stores the labels in the `Survey` instead of in the data. We can recreate this as follows:

```{r}
sav_cleaned <- from_labelled(sav)
attr(sav_cleaned, "labels") <- NULL

all.equal(srv$data, sav_cleaned)
```

The last line shows that `sav_cleaned` is equal to the data stored in the survey.

#### Using `$` with a survey

Because a `Survey` is implemented in R6, the `$` operator is reserved for accessing the contents (seen above) and methods (explained below) of the `Survey`, and not columns in the data directly:

```{r}
# Fail
srv$q1

# Success
head(srv$data$q1, n = 1L)
```

#### Object oriented

`R6` is a packages which helps us do object orientation in R. In practice this means that many functions are defined inside the `Survey` itself, and can be called using `$`. For instance, all **get** and **set** functions are defined inside the `Survey` and have S3 wrappers:

```{r}
# S3 method for setting a label
srv <- set_label(srv, StartDate = "Date when respondent opened the Survey.")

# R6 method
srv$set_label(StartDate = "Timestamp when respondent opened the Survey.")
```

Both of these functions do the same thing, because the S3 method is just a function that calls the R6 method. However, the S3 method returns a copy of the Survey you passed to the function, while the R6 method modifies the existing survey (in place). 

#### Mutable

R6 objects are mutable, and since the `Survey` is built on R6, it is also mutable. This means that we can make changes to the `Survey` without having to assign (`<-`) the result, like you saw an example of above (R6 method for setting label). Because it is mutable, this allows the `Survey` to have hidden fields that behave like the `data.table` itself.


However, mutable objects are rather atypical for R objects, and so I have made the following decision in order to make methods more predictable: 

- All S3 methods (e.g. set_label()) return a copy of the Survey, and must be assigned (`<-`).
- All R6 methods (accessed with `$`) modify the Survey in place.

## Detailed usage

As we can see, simply calling `survey()` on the data will create a `Survey_tbl`, this is because `haven` returns a `tbl_df` by default. If we instead wanted a  `Survey_dt` (i.e., use `data.table`), we have two options:

```{r}
dtm <- survey(data.table::as.data.table(sav)) # Option 1: Manual.
dt <- survey_dt(sav) # Option 2: Use the Survey_dt constructor.
"Survey_dt" %in% class(dt) && "Survey_dt" %in% class(dtm)
```

That's it. We can either call the generic constructor `survey()`, or a specific  one which coerces the input data, these are: 

- `survey_df()` 
- `survey_dt()` 
- `survey_tbl()`, requires [dplyr](https://github.com/hadley/dplyr).

#### `[` and `[[`

When working with surveys, we can use `[` and `[[` just like on a regular `data.frame`:
```{r}
df <- survey_df(sav)
df[["test"]] <- "test"
df[1L, "test"]
```

Or `data.table`: 
```{r}
dt <- survey_dt(sav)
head(dt[, test := "test"][1L, test])
```

When subsetting a `Survey`, the new object will also be a `Survey`:
```{r}
dt_slice <- dt[1:5L, .(test)]
class(dt_slice)
```

But only if the operation returns an object that inherits from `data.frame`:
```{r}
class(dt_slice[, test])
```

Overall, a `Survey` should behave like the underlying data. The only exception to this, is that `$` is reserved for accessing public fields in R6 objects:
```{r}
identical(df$test, df[["test"]]) # FALSE
```

Because the `Survey` class is implemented in R6, they are mutable and allows the  `Survey_dt` subclass to have hidden fields behave similar to the `data.table` itself.

#### Hidden fields

As mentioned previously, the `Survey` lets us keep track of extra information to automate certain processes - these are kept in the following hidden fields:

- **labels**: The question text for each variable.
- **associations**: Associations, such as latent association in the case of PLS-PM.
- **marketshares**: The marketshare for each entity in the study.
- **config**: Name of study, segment, missing % cutoff etc.
- **translations**: A dictionary of common words and their replacement. Used in tables and plots.

To retrieve one or more labels, you can use `get_label()`:
```{r}
# Get the label for Q1 using the R6 method
q1_label <- dt$get_label("q1")

# S3 Method (recommended). Creates a copy and must be assigned.
q1_label <- get_label(dt, "q1")

# To get all labels, just drop the second argument
all_labels <- get_label(dt)
```

The object `dt` already contains labels in this case, because `survey()` has found them after reading in the SPSS-file in `seamless`. You can also set one or more labels manually using `set_label()`:

```{r}
# Get the label for Q1 using the R6 method
dt$set_label(q16 = "Ideal")

# S3 Method (recommended). Creates a copy and must be assigned.
dt <- set_label(dt, q6 = "Vs expectations")

# To set several labels, use the list argument.
dt <- set_label(dt, list = list(q1 = "This is the mainentity", q3 = "Overall"))
```

This is more or less the pattern used to *get* or *set* all hidden fields in 
a `Survey`.

Get hidden fields:

- `get_label()`
- `get_association()`
- `get_marketshare()`
- `get_config()`
- `get_translation()`

Set hidden fields:

- `set_label()`
- `set_association()`
- `set_marketshare()`
- `set_config()`
- `set_translation()`

Having set the labels above, the Survey will keep track of them as the data changes:

```{r}
dt_subset <- dt[, .(q3, q6, q16)]
names(dt_subset) <- paste0("example", 1:3L)
get_label(dt_subset) # No second argument, all labels are returned.
```

#### Model and entities

You can also use the function `model()` to return a list of variables in the data, their names, type and label:

```{r}
# R6 method
model_example <- dt_subset$model()

# S3 (recommended)
model(dt_subset)
```

If we set associations, these are also marked next to the variable type in the output from `model()`:
```{r}
dt_subset <- set_association(dt_subset, epsi = "example1")
model(dt_subset)
```

You can also use `common = TRUE` to guess associations based on variable names:

```{r}
dt <- set_association(dt, common = TRUE)
# Call model(dt) to see which variables have been identified.
get_association(dt, "mainentity")
```

After identifying common associations and having Q1 identified as our mainentity variable, we can use `entities()` to get a summary: 

```{r}
dt$entities()
```

If we set a marketshare, rename to "percent_missing" and set cutoff in config:

```{r}
names(dt)[names(dt) == "Andel_Missing"] <- "percent_missing"
dt <- set_marketshare(dt, reporttool = 1L)
dt <- set_config(dt, cutoff = .1) # 10%
dt$entities()
```

## More information

[Introduction](https://github.com/itsdalmo/reporttoolDT/blob/master/vignettes/introduction.Rmd): 
```{r, eval = FALSE}
vignette("introduction", package = "reporttoolDT")
```

[Preparing data](https://github.com/itsdalmo/reporttoolDT/blob/master/vignettes/prepare.Rmd):
```{r, eval = FALSE}
vignette("prepare", package = "reporttoolDT")
```

[Other functions](https://github.com/itsdalmo/reporttoolDT/blob/master/vignettes/other.Rmd): 
```{r, eval = FALSE}
vignette("other", package = "reporttoolDT")
```
